---
title: "Homework 1"
author: "Laya Buchanan"

date: 2020-09-21
output: github_document
---

This is my submission for the second homework assignment for P8105.  


```{r message = FALSE, echo = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```


# Problem 1: Mr Trash Wheel


```{r}

```

# Problem 2: NYC Transit Data

## Read in some data

Read in the NYC Transit dataset. 

```{r message = FALSE, echo = FALSE}
transit_df = read_csv("./data/NYC_Transit_Data.csv")
```

## Clean the data

Clean up column names:

```{r warning = FALSE, message = FALSE, results = "hide"}
transit_df = janitor::clean_names(transit_df)
```

Select relevant variables and convert the entry variable from character to logical:
```{r message = FALSE, results = "hide"}
transit_df =
  select( transit_df, line, station_name, station_latitude, station_longitude, route1, route3, route5, route7, route9, route11, entry, entrance_type, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))

```

```{r}

```


# Problem 3: FiveThirtyEight Data

## Read in some data

Read in the the FiveThirtyEight data

```{r message = FALSE, results = "hide"}
pols_month = read_csv("./data/pols_month.csv")
snp = read_csv("./data/snp.csv")
unemployment = read_csv("./data/unemployment.csv")
```

## Clean the data

Clean up column names:

```{r message = FALSE, results = "hide"}
pols_month = janitor::clean_names(pols_month)
snp = janitor::clean_names(snp)
unemployment = janitor::clean_names(unemployment)
```